---
title: "FORS 505"
author: "Colin Kirkmire"
date: "April 28, 2016"
output: word_document
---

(4) Compute a 90% CI for the number of TPHA on the whole tract
```{r}
A_1<-c(10.3,10.3,11.2,13.6)
A_2<-c(11.2,8.5,9.4,6.8)
A_3<-vector(mode="numeric", length=0)
A_4<-c(9.8,8.9,12.2,5.6,3.4)
B_1<-c(18.9,14.5,12.9,14.2,9.9)
B_2<-c(11.2,21.3,14.2,15.3,8.2)
B_3<-c(7.6)
B_4<-c(10.3,17.4,19.2)
B_5<-c(3.4,13.5,19.5,18.5,14.6)
```

```{r}
#Estimate from Strata A#
total_tree_function <- function(x) {
   result<-18*10000*(length(x)/25)
   return(result)}

T_hat_A1<-total_tree_function(A_1)
T_hat_A2<-total_tree_function(A_2)
T_hat_A3<-total_tree_function(A_3)
T_hat_A4<-total_tree_function(A_4)
all_plots_A<-c(T_hat_A1,T_hat_A2,T_hat_A3,T_hat_A4)
est_tree_total_A<-mean(all_plots_A)
var_tree_total_A<-(1/4)*(var(all_plots_A))
```

```{r}
#Estimate from Strata B#
BAF<-4

tf <- function(x) {
   result<-1/(3.14*((x/200)^2))
   return(result)}
T_hat_B1<-42*BAF*(sum(tf(B_1)))
T_hat_B2<-42*BAF*(sum(tf(B_2)))
T_hat_B3<-42*BAF*(sum(tf(B_3)))
T_hat_B4<-42*BAF*(sum(tf(B_4)))
T_hat_B5<-42*BAF*(sum(tf(B_5)))
all_plots_B<-c(T_hat_B1,T_hat_B2,T_hat_B3,T_hat_B4,T_hat_B5)
est_tree_total_B<-mean(all_plots_B)
var_tree_total_B<-(1/5)*(var(all_plots_B))
```

```{r}
tpha_estimate<-(sum(est_tree_total_A,est_tree_total_B))/60
SE4<-(sqrt(var_tree_total_A+var_tree_total_B))/60

#90 CI for number of  tpHA#
tpha_estimate + c(-1,1)* SE4 * qt(p=.95,df=7)
```

(5) 90% CI for baHA

```{r}
#Estimate from Strata A#
ba_total_function_A <- function(x) {
   result<-sum(18*400*(((((x)/100)/2)^2)*3.14))
   return(result)}
Tba_hat_A1<-ba_total_function_A(A_1)
Tba_hat_A2<-ba_total_function_A(A_2)
Tba_hat_A3<-ba_total_function_A(A_3)
Tba_hat_A4<-ba_total_function_A(A_4)

Tba_all_plots_A<-c(Tba_hat_A1,Tba_hat_A2,Tba_hat_A3,Tba_hat_A4)
est_ba_total_A<-mean(Tba_all_plots_A)
var_ba_total_A<-(1/4)*(var(Tba_all_plots_A))

#Estimate from Strata B#
ba_total_function_B <- function(x) {
   result<-BAF*length(x)*42
   return(result)}
Tba_hat_B1<-ba_total_function_B(B_1)
Tba_hat_B2<-ba_total_function_B(B_2)
Tba_hat_B3<-ba_total_function_B(B_3)    
Tba_hat_B4<-ba_total_function_B(B_4)
Tba_hat_B5<-ba_total_function_B(B_5)

Tba_all_plots_B<-c(Tba_hat_B1,Tba_hat_B2,Tba_hat_B3,Tba_hat_B4,Tba_hat_B5)
est_ba_total_B<-mean(Tba_all_plots_B)
var_ba_total_B<-(1/5)*(var(Tba_all_plots_B))
```

```{r}
bapHA_estimate<-(sum(est_ba_total_A,est_ba_total_B))/60
SE5<-(sqrt(var_ba_total_A+var_ba_total_B))/60

#90 CI for number of  bapHA#
bapHA_estimate + c(-1,1)* SE5 * qt(p=.95,df=7)
```

(6) Estimate the total ba in trees larger than 10cm DBH as well as its standard error.
```{r}
#BA>10cm for strata A#
Tba_hat_A1_ten<-ba_total_function_A(A_1[A_1>10])
Tba_hat_A2_ten<-ba_total_function_A(A_2[A_2>10])
Tba_hat_A3_ten<-ba_total_function_A(A_3[A_3>10])
Tba_hat_A4_ten<-ba_total_function_A(A_4[A_4>10])

Tba_all_plots_A_ten<-c(Tba_hat_A1_ten,Tba_hat_A2_ten,Tba_hat_A3_ten,Tba_hat_A4_ten)
  
est_ba_total_A_ten<-mean(Tba_all_plots_A_ten)
var_ba_total_A_ten<-(1/4)*(var(Tba_all_plots_A_ten))

#BA>10cm for strata B#
Tba_hat_B1_ten<-ba_total_function_B(B_1[B_1>10])
Tba_hat_B2_ten<-ba_total_function_B(B_2[B_2>10])
Tba_hat_B3_ten<-ba_total_function_B(B_3[B_3>10])    
Tba_hat_B4_ten<-ba_total_function_B(B_4[B_4>10])
Tba_hat_B5_ten<-ba_total_function_B(B_5[B_5>10])

Tba_all_plots_B_ten<-c(Tba_hat_B1_ten,Tba_hat_B2_ten,Tba_hat_B3_ten,Tba_hat_B4_ten,Tba_hat_B5_ten)
  
var_ba_total_B_ten<-(1/5)*(var(Tba_all_plots_B_ten))

est_ba_total_B_ten<-mean(Tba_all_plots_B_ten)

#Estimated Total BA in meters for whole stand#
ba_estimate_ten<-(sum(est_ba_total_A_ten,est_ba_total_B_ten))
ba_estimate_ten

SE6<-sqrt(var_ba_total_A_ten+var_ba_total_B_ten)
SE6
```


(7) Estimate the average diameter of trees in stratum A and, seperately, of trees in stratum B.

No, you can not use the same estimators for each stratum because stratum B trees were selected using an unequal probability sampling method (variable radius plots) in which the larger trees were selected with a preference.

```{r}
#Average diameter in stratum A#
total_diam_function <- function(x) {
   result<-18*10000*sum(x)/25
   return(result)}

diam_A1<-total_diam_function(A_1)
diam_A2<-total_diam_function(A_2)
diam_A3<-total_diam_function(A_3)
diam_A4<-total_diam_function(A_4)

est_diam_total_A<-(sum(diam_A1,diam_A2,diam_A3,diam_A4))/4

est_average_diam_A<-est_diam_total_A/est_tree_total_A
est_average_diam_A

#Average diameter in stratum B#
total_diam <- function(x) {
   result<-BAF*42*(sum(x/(3.14*(x/200)^2)))
   return(result)}

diam_B1<-total_diam(B_1)
diam_B2<-total_diam(B_2)
diam_B3<-total_diam(B_3)    
diam_B4<-total_diam(B_4)
diam_B5<-total_diam(B_5)

est_diam_total_all_plots_B<-(sum(diam_B1,diam_B2,diam_B3,diam_B4,diam_B5))/5
est_average_diam_B<-est_diam_total_all_plots_B/est_tree_total_B
est_average_diam_B
```










