---
title: 'HW #6'
author: "Colin Kirkmire"
date: "April 9, 2016"
output: html_document
---


(a)
```{r}
sat <- read.csv("C:/Users/Colin/Desktop/Rprojects/Stat452/sat.csv")

#Relationship between SAT,Takers and Ranks#
rel.sat<-sat[,c('SAT','Takers','Rank')]

library(GGally)
ggpairs(rel.sat)
```

There is a strong, negative relationship between the Takers variable and SAT(r=-.86). However, the Taker's variable appears to be decrease exponentially as statewide average SAT increases. As the percent of high school students who took the SAT increases, the state's average total score decreases roughly exponentially. 

In contrast, there is a strong positive linear relationship between Rank and SAT (r=.88).  The state's average total SAT score increases as the median percentile rank of test-takers increases. 

The relationship between rank and takers appears to be very strong (r=-.94) and negatively linear. When the percent of a state's high schoolers who take the SAT increases, the median percentile rank of test-takers decreases.  

The three states with the lowest SAT scores (South Carolina, Georgia and North Carolina) are apparent outliers in all plots of relationships that include SAT scores.


(b)
```{r}
sat.lm<-lm(sat$SAT~sat$Takers+sat$Rank)

#Resid vs Fitted#
plot(sat.lm,which=c(1))

#Normality of Residuals#
hist(sat.lm$residuals)

#Cooks Distance#
plot(sat.lm,which=c(4))

#leverage#
lev<-hat(model.matrix(sat.lm))

plot(lev)

#Added Variable Plots#
library(car)
avPlots(sat.lm)

#Component plus resid plots#
cr.bplots<-crPlots(sat.lm)

```
The residuals vs fitted values plot reveals a lack of homoscedasticity. There is a pattern in  which both lower and higher fitted values have high (positive) residuals and intermediate fitted values have lower (more negative) residuals.

The residuals display a lack of normality as they appear to be skewed moderately to the left in the histogram of residuals.

Iowa, South Carolina and Connecticut have very high Cook's distance values. This indicates that these states may be heavily influencing the model. 

The added variable plots do not reveal any high leverage points but there are several notable low outliers in both plots that are from the low SAT scores in states including South Carolina, North Carolina and Georgia.

The component plus residual plot of average SAT scores vs Takers reveals non-linearity.  This suggests that that the takers variable should be transformed.  This is consistant with the non-linear (negative exponential) relationship that was observed between SAT and Takers in part A.

(c)
```{r}
log.sat.lm<-lm(sat$SAT~log(sat$Takers)+sat$Rank)

#Resid vs Fitted#
plot(log.sat.lm,which=c(1))

#Normality of Residuals#
hist(log.sat.lm$residuals)

#Cooks Distance#
plot(log.sat.lm,which=c(4))

#leverage#
lev.c<-hat(model.matrix(log.sat.lm))

plot(lev.c)

#Added Variable Plots#
library(car)
avPlots(log.sat.lm)

#Component plus resid plots#
crPlots(log.sat.lm)
```
The homoscedasticity of the residuals is improved and the pattern described in the residual vs fitted plot of part B is no longer evident. 

The normality of the residuals is slightly improved but there is still a left skew due to several very large negative residuals.


Two of these very large negative residuals were also apparent in the first model and belong to North Carolina and South Carolina. 


These two states also appear in the plot of Cook's distance as having a great influence on the model.  Mississippi also exhibits a high Cook's distance.

The states that appeared in the untransformed added variable plots are still present as outliers in the new added variable plots.

However, the nonlinearity observed in the untransformed component plus residual plot of SAT vs takers has been corrected with the log transformation of takers.


(d)
```{r}
#Fit model with log(Takers) and Rank (no interaction)#

logt.sat.lm<-lm(sat$SAT~log(sat$Takers)+sat$Rank)

sat$resid<-(resid(logt.sat.lm))

ordered.sat<-sat[order(logt.sat.lm$resid),]

```
South Carolina,North Carolina and Mississippi performed worst after adjusting for the number and quality of students taking the test.  The top performing states were New Hampshire, Iowa and Connecticut.


(e)
```{r}
#Relationship between Expend, Years and Income#
rel.expend.years.income<-sat[,c('Years','Expend','Income','resid','State')]

library(GGally)
ggpairs(rel.expend.years.income[,1:4])

mean(rel.expend.years.income$Expend)

sd(rel.expend.years.income$Expend)


```
The correlation coeficient of income vs residual (r=.135) demonstrates that there is a lack of linear relationship between these variables.  This suggests that median family income is not related to the residual.  However, Years and Expenditure  appear to have stronger positive linear relationships with the residuals albeit these are not very strong relationships  (r=.41 and r=.50 respectively).  

Years is not stongly correlated with expenditure (r=.0598) or income (r=.135).  This means that both the amount that a state spends per student as well as median test taker family income are not well correlated with how many years of formal study that sudents had in social sciences, natural sciences and humanities.

There is an obvious outlier in terms of spending per student in secondary education (Expend).  Alaska spends over $5,000 per secondary school student. The mean and standard deviation of expenditure is $2,297 and $614, respectively. 

There are also two states that are outlying in regard to median family income vs both years and expend. These states (Alaska and Louisiana) also happen to be the highest states overall in terms of median family income of test-takers.


(f)
```{r}

#Fit model with log(Takers) model w Expend, Income and Years#

f.logt.sat.lm<-lm(SAT~log(Takers)+Rank+Expend+Years, data=sat)

#Resid vs Fitted#
plot(f.logt.sat.lm,which=c(1))

#Normality of Residuals#
hist(f.logt.sat.lm$residuals)

#Resid vs Fitted#
plot(f.logt.sat.lm,which=c(1))

#Cooks Distance#
plot(f.logt.sat.lm,which=c(4))

#leverage#
lev<-hat(model.matrix(f.logt.sat.lm))
plot(lev)

#Added Variable Plots#
library(car)
avPlots(f.logt.sat.lm)

#Component plus resid plots#
crPlots(f.logt.sat.lm)

```

The variance of residuals appears constant accross fitted values. Although there are several large positive and negative residuals they are spread accross fitted values.

The normality of the residuals is better than the previous normality but there remains a slight left skew due to a very large negative residual.


Alaska has a Cook's distance that dwarfs that of the other states.  This indicates that Alaska is highly influencing the model.

The added variable plots have no obvious high leverage points or outliers with the exception of the SAT vs expenditure av plot which shows that Alaska has a lot of leverage in the plot. The added variable plots reveals where Alaska's influence in the model is from.


All compenent plus residual plots display acceptable linear relationships with the exception of the plot of the plot of component+resid(SAT) vs Expenditure. This nonlinear relationship can be attributed to a single high leverage point that represents a state that spends a much larger amount per high school student than the other states: Alaska.


Alaska has the highest expenditure per student by a large amount.  Anecdotely, I have heard that much of the money spent on high school students in Alaska involves transporting them to various extracurricular events. Bringing teams together for athletic competitions may involve flying the entire basketball team to another location since the state is so vast and schools are often otherwise virtually inaccesible.

The increase in money that the other states spend per student likely relates better to improved SAT scores because more funds are allocated towards textbooks, teachers, etc. 

Alaska's high state ependiture per student due to travel cost will likely be of little use in a model for state SAT scores and will only serve to "pull" the multiple regression towards Alaska since it has such a high Cook's distance. Therefore, it is justified to remove Alaska from the model.

(g)
```{r}
#fit w/o Alaska#
sat.sansAK<-sat[!sat$State==('Alaska'),]

sansAK.lm<-lm(SAT~log(Takers)+Rank+Expend+Years+Income, data=sat.sansAK)

#Refitting model from Part D#
logt.sansAK.lm<-lm(SAT~log(Takers)+Rank, data=sat.sansAK)

#Partial F-test#
anova(sansAK.lm,logt.sansAK.lm)

```
The large F-statistic provides strong evidence that at least one of the variables added in the model that includes the three variables (Expend, Years, and Income) results in a statistically different model.


(h)
```{r}
sansAK.sum<-summary(sansAK.lm)

#Confidence intervals for coefficients#

confint(sansAK.lm,level=.95)

```

Takers: 
We are 95% confident that a one percent increase in the percent of high school students in the state who took the SAT is associated with a decrease in state average SAT score between .034 and .586 given the other predictor variables included in the model are held constant.

Rank:
Given all other predictors included are held constant, we are 95% confident that the average state SAT score increases between 1.01 and 9.11 points for every unit increase in the median percentile rank of the test-takers within their schools.

Expend:
We are 95% confident that for every hundred dollar increase in per studen expenditure, the average SAT increases by between 2.31 and 5.43 points (given all other predictor variables constant).

Years:
We are 95% confident that the state average SAT score increases between 2.43 and 24.56 points for ever additional average number of years that test takers had in social sciences, natural sciences and humanities (given all other predictor variables constant).

Income:
We are 95% confident that the state average SAT score increases between -.11 and .34 points for every additional hundred dollars in median family income (given all other predictor variables constant).  The increase in median family income may have a negative, zero, or positive difference in state average SAT score given all other predictor variables are held constant.

(i)
```{r}
red.sat.lm<-lm(SAT~log(Takers)+Rank+Expend,data=sat.sansAK)

woIncome.sum<-summary(red.sat.lm)

#CI for Coefficients#
confint(red.sat.lm,level=.95)

```
The coeffient of expenditure in the reduced model no longer can be interpreted as the increase in state average SAT score given rank, takers, years and income are held constant. Rather, the coefficient of expenditure only reflects the estimated increase in state average SAT score given takers, rank and expenditure are held constant. The coefficient of expenditure became larger because it absorbed some of the variability in state average SAT score that was previously attributed to median family income.




(j)
```{r}
library(robustbase)


robust.lm<-lmrob(SAT~log(Takers)+Rank+Expend+Years+Income, data=sat)

confint(robust.lm,level=.95)
confint(sansAK.lm)


robust.lm$rweights[29]

```
The confidence intervals of coefficients between the two methods of mlr are quite similar.  However, it was striking how much narrower the CI for the intercept was for the robust regression. The robust.lm log(takers) CI was also narrower than that of the initial model.  Interstingly, all of the remaining variables (Rank, Expend, years, income) had intial model 95% CI that could fit within those of their robust.lm counterparts.


A weight of zero was assigned to Alaska by the robust regression function.  The outlier therefore did not contribute to the robust regression model.





















