---
title: 'HW #6'
author: "Colin Kirkmire"
date: "April 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
 sat <- read.csv("C:/Users/Colin/Desktop/Rprojects/Stat452/sat.csv")

sat.lm<-lm(sat$SAT~sat$Takers+sat$Rank)

#Check residuals#

plot(resid(sat.lm)~predict(sat.lm))
hist(resid(sat.lm),col="gray")
qqnorm(resid(sat.lm))

#Try log transformation#
plot(sat$SAT~log(sat$Takers))
plot(sat$SAT~(sat$Rank))

sq.sat.lm<-lm(sat$SAT~log(sat$Takers)+sat$Rank^2)


#Re-Check residuals#

plot(resid(sq.sat.lm)~predict(sq.sat.lm))
hist(resid(sq.sat.lm),col="gray")
qqnorm(resid(sq.sat.lm))


summary(sat.lm)
```

(a)
```{r}
#Relationship between SAT,Takers and Ranks#
rel.sat<-sat[,c('SAT','Takers','Rank')]

library(GGally)
ggpairs(rel.sat)
```

There is a strong, negative relationship between the Takers variable and SAT. However, the Taker's variable appears to be decrease exponentially as statewide average SAT increases.
As the percent of high school students who took the SAT increases, the state's average total score decreases exponentially. 

In contrast, there is a strong positive  relationship between Rank and SAT.  The state's average total SAT score increases as the median percentile rank of test-takers increases. 

The relationship between rank and takers appears to be very strong (r=-.94) and negatively linear. When the percent of a state's high schoolers who take the SAT increases, the median percentile rank of test-takers decreases.  

##CHECK for outliers##

(b)
```{r}
b.sat.lm<-lm(sat$SAT~sat$Takers+sat$Rank)

plot(resid(b.sat.lm)~b.sat.lm$fitted.values)

library(car)
crPlots(b.sat.lm)

```

(c)
```{r}
log.sat.lm<-lm(sat$SAT~log(sat$Takers)+sat$Rank)

plot(resid(log.sat.lm)~log.sat.lm$fitted.values)

library(car)
crPlots(log.sat.lm)
```

(d)
```{r}
#Fit model with log(Takers) and Rank (no interaction)#

logt.sat.lm<-lm(sat$SAT~log(sat$Takers)+sat$Rank)

sat$resid<-(resid(logt.sat.lm))

ordered.sat<-sat[order(logt.sat$resid),]

```
After adjusting for the number and quality of students taking the test, South Carolina,



(e)
```{r}

```

(f)
```{r}

```

(g)
```{r}

```

(h)
```{r}

```

(j)
```{r}

```






