---
title: 'HW #6'
author: "Colin Kirkmire"
date: "April 9, 2016"
output: html_document
---


(a)
```{r}
sat <- read.csv("C:/Users/Colin/Desktop/Rprojects/Stat452/sat.csv")

#Relationship between SAT,Takers and Ranks#
rel.sat<-sat[,c('SAT','Takers','Rank')]

library(GGally)
ggpairs(rel.sat)
```

There is a strong, negative relationship between the Takers variable and SAT(r=-.86). However, the Taker's variable appears to be decrease exponentially as statewide average SAT increases. As the percent of high school students who took the SAT increases, the state's average total score decreases roughly exponentially. 

In contrast, there is a strong positive linear relationship between Rank and SAT (r=.88).  The state's average total SAT score increases as the median percentile rank of test-takers increases. 

The relationship between rank and takers appears to be very strong (r=-.94) and negatively linear. When the percent of a state's high schoolers who take the SAT increases, the median percentile rank of test-takers decreases.  

The three states with the lowest SAT scores (South Carolina, Georgia and North Carolina) are apparent outliers in all plots of relationships that include SAT scores.


(b)
```{r}
sat.lm<-lm(sat$SAT~sat$Takers+sat$Rank)

#Resid vs Fitted#
plot(sat.lm,which=c(1))

#Normality of Residuals#
hist(sat.lm$residuals)

#Cooks Distance#
plot(sat.lm,which=c(4))

#leverage#
lev<-hat(model.matrix(sat.lm))

plot(lev)

#Component plus resid plots#
library(car)
cr.bplots<-crPlots(sat.lm)

identify(cr.bplots)
```
The residuals vs fitted values plot reveals a lack of homoscedasticity in the residuals. There is a pattern in  which both lower and higher fitted values have high (positive) residuals and intermediate fitted values have lower (more negative) residuals.

The residuals display a lack of normality as they appear to be skewed moderately to the left in the histogram of residuals.

Iowa, South Carolina and Connecticut have very high Cook's distance values. This indicates that these states may be heavily influencing the model. 

The component plus residual plot of average SAT scores vs Takers reveals non-linearity.  This suggests that that the takers variable should be transformed.  This is consistant with the non-linear (negative exponential) relationship that was observed between SAT and Takers in part A.

(c)
```{r}
log.sat.lm<-lm(sat$SAT~log(sat$Takers)+sat$Rank)

#Resid vs Fitted#
plot(log.sat.lm,which=c(1))


#Cooks Distance#
plot(log.sat.lm,which=c(4))

#leverage#
lev.c<-hat(model.matrix(log.sat.lm))

plot(lev.c)

#Component plus resid plots#
library(car)
crPlots(log.sat.lm)
```
The homoscedasticity of the residuals is improved and the pattern described in part B is no longer evident. However, there is is still a very large negative residual that was also apparent in the first model that belongs to South Carolina. This state also appears in the plot of Cook's distance as having a great influence on the model.  There are two additional states that exhibit high Cook's distances, Mississippi and North Carolina.


(d)
```{r}
#Fit model with log(Takers) and Rank (no interaction)#

logt.sat.lm<-lm(sat$SAT~log(sat$Takers)+sat$Rank)

sat$resid<-(resid(logt.sat.lm))

ordered.sat<-sat[order(logt.sat.lm$resid),]

```
South Carolina,North Carolina and Mississippi performed worst after adjusting for the number and quality of students taking the test.  The top performing states were New Hampshire, Iowa and Connecticut.


(e)
```{r}
#Relationship between Expend, Years and Income#
rel.expend.years.income<-sat[,c('Years','Expend','Income','resid','State')]

library(GGally)
ggpairs(rel.expend.years.income[,1:4])

mean(rel.expend.years.income$Expend)

sd(rel.expend.years.income$Expend)


```
The correlation coeficient of income vs residual (.135) suggests that there is a lack of linear relationship between these variables.  This suggests that median family income is not related to the residual.  However, Years and Expenditure  appear to have stronger positive linear relationships with the residuals albeit these are not very strong relationships  (.41 and .50 respectively).  

Years is not stongly correlated with expenditure (r=.0598) or income (r=.135).  This means that both the amount that a state spends per student as well as median test taker family income are not correlated with how many years of formal study that sudents had in social sciences, natural sciences and humanities.

There is an obvious outlier in terms of spending per student in secondary education (Expend).  Alaska spends over $5,000 per student per student. The mean and standard deviation of expenditure is $2,297 and $614, respectively. 

There are also two states that are outlying in regard to median family income vs both years and expend. These states (Alaska and Louisiana) also happen to be the highest states overall in terms of median family income of test-takers.


(f)
```{r}

#Fit model with log(Takers) model w Expend, Income and Years#

f.logt.sat.lm<-lm(SAT~log(Takers)+Rank+Expend+Years, data=sat)


#Resid vs Fitted#
plot(f.logt.sat.lm,which=c(1))

#Cooks Distance#
plot(f.logt.sat.lm,which=c(4))

#leverage#
lev<-hat(model.matrix(f.logt.sat.lm))

plot(lev)


library(car)
crPlots(f.logt.sat.lm)


```
Alaska has the highest expenditure per student by a large amount.  Anecdotely, I have heard that much of the money spent on high school students in ALaska involves transporting them to various extracurricular events. Bringing teams together for athletic competitions may involve flying the entire basketball team to another location since the state is so vast and schools are often otherwise virtually inaccesible.

This increase in state ependiture per student for travel cost will likely be of little use in a model for state SAT scores and will only serve to "pull" the least squares reggression line towards it since it has such a high leverage. Therefore it is justified to remove Alaska from the model.

(g)
```{r}
#fit w/o Alaska#
sat.sansAK<-sat[!sat$State==('Alaska'),]

sansAK.lm<-lm(SAT~log(Takers)+Rank+Expend+Years+Income, data=sat.sansAK)

#Refitting model from Part D#
logt.sansAK.lm<-lm(SAT~log(Takers)+Rank, data=sat.sansAK)

#Partial F-test#
anova(sansAK.lm,logt.sansAK.lm)

```
The large F-statistic provides strong evidence that at least one of the variables added in the model that includes the three variables (Expend, Years, and Income) results in a statistically different model.


(h)
```{r}
sansAK.sum<-summary(sansAK.lm)

#Confidence intervals for coefficients#

confint(sansAK.lm,level=.95)

```

(i)
```{r}
red.sat.lm<-lm(SAT~log(Takers)+Rank+Expend,data=sat.sansAK)

woIncome.sum<-summary(red.sat.lm)

#Expend CI#
woIncome.sum$coefficients[4]+c(-1,1)*woIncome.sum$coefficients[4,2]*woIncome.sum$coefficients[4,3]

```
ans=given income vs not given income


(j)
```{r}
library(robustbase)


robust.lm<-lmrob(SAT~log(Takers)+Rank+Expend+Years+Income, data=sat)

confint(robust.lm,level=.95)

robust.lm$rweights[29]

```
A weight of zero was assigned to Alaska by the robust regression function.  The outlier therefore did not contribute to the robust regression model.





















