---
title: "Lab#7"
author: "Colin Kirkmire"
date: "March 21, 2016"
output: word_document
---

Homework [1]:
```{r setup, include=FALSE}
egyptians <- read.csv("F:/STAT 458/Lab7/egyptians.csv")

model1 <- lm(MB ~ BH + BL +NH, data=egyptians)
model1
summary(model1)
plot(fitted(model1), resid(model1), main="residual plot",
xlab="predicted response",
ylab="residuals")
abline(h=0, lty=2)
var1 <- predict(model1)
var2 <- fitted (model1)
```

Homework[2]: Using the basscatch.csv data set, pick one or 2 variables which you hypothesize might be 0 and use this subset F test procedure to gather evidence on
your hypothesis.

```{r}
basscatch <- read.csv("F:/STAT 458/Lab7/basscatch.csv")

#Pairs Plots
library(GGally)
ggpairs(basscatch)

#Subset F test
modelfull <- lm(catch ~ residence + size + access + structure)
modelpart <- lm(catch ~ residence + size)
summary(modelfull)
anova(modelfull)
summary(modelpart)
anova(modelpart)
```

Homework[2]: The data set crops.csv is mercury poisoning results of an agricultural experiment involving 3 kinds of crops (corn, wheat, and barley) planted in mercury tainted soil (called sludge). There are 6 levels of soil contamination, and the
mercury content contained in the plants (the response variable) is measured. Compare corn with wheat to see if the β3 (the interaction term) is significant.
Possibly construct a scatter plot with the 2 ablines() superimposed on the plot to see what you will end up with on the hypothesis test. Use the α = 0.05 level. You may want to add an x2 column of 0's and 1's as I did in my demonstration. Remember you will only be using rows 1 through 60 (i.e., [1:60]) for this test.

```{r}
crops <- read.csv("F:/STAT 458/Lab7/crops.csv")

crops<-crops[1:60,]
```



