---
title: "Lab#9"
author: "Colin Kirkmire"
date: "April 7, 2016"
output: word_document
---

Homework [1]: Using your full model from last week's babies data from the UsingR package, find a reduced model using AIC analysis and compare with your model reduced using partial-F tests.

```{r}
library("UsingR")
babies<-babies

notthese<-(babies$gestation==999 | (babies$age==99)) | (babies$inc==98)
tmp<-babies[!notthese,1:23]

babywt.lm<-lm(wt~gestation+ht+number+time+wt1+parity+date+age,data=tmp)

library(MASS)
stepAIC(babywt.lm)
```
The results of the step AIC matches the results of the partial F test exactly. The step AIC suggests the removal of only the age predictor, just as the partial F test shows that the removal of the age predictor is not statistically significant.  


Homework [2]: Take some quality time and analyze the example problem as well as the listing of techniques and R commands, pondering approaches, techniques and conclusions. Then write about 50 words saying what might be useful or not useful for you in what you see.


The cor() and pairs() commands will be very useful in exploring relationships between all variables. The partial F test of linear models provides a comparison between models with a removed term.  

The interaction plots will helpful in determining whether the effect of one independent variable depends on the level of another independent variable.  

The step AIC command will be useful in model building, however, it is important not to rely on the results of the stepwise regression.

Of course, the residual() and qqplot/qqnorm() commands are critical for assesing the assumptions of the linear model.

Homework [3]: You did an interaction plot in lab 7 on factors of the ratanxiety.csv data. Redo the interaction plot using the command interaction.plot() and
compare the results.

```{r}
ratanxiety <- read.csv("E:/STAT 458/Lab7/ratanxiety.csv")

interaction.plot(ratanxiety$dose, ratanxiety$product, ratanxiety$anxiety, legend=TRUE)
```

