---
title: "Lab11"
author: "Colin Kirkmire"
date: "April 24, 2016"
output: word_document
---


Homework [1]: The gestation data set (described below has factors related to 96 mammal species, their dimensions, and birthing times. This data set is in the Sleuth3 package, listed as case0902. II made some descriptives of the information and found that transforming all variables to natural logs helped in modeling. Use the information provided below and generate your best model of this information, using the various multiregression techniques you have learned. Present about a 50 to 100 word defense of your resulting best model. We want brainweight (or more properly log(brainweight)) as our response variable, and include any interaction terms you feel might be important for your finished model.

```{r}
library(Sleuth3)

animals<-as.data.frame(case0902)

#Pairs Plots#
library(GGally)
plot(log(animals$Brain)~animals$Body)
plot(log(animals$Brain)~animals$Gestation)
plot(log(animals$Brain)~animals$Litter)
plot(animals$Body~animals$Gestation)
plot(animals$Body~animals$Litter)
```

It appears as thought the linearity of the relationship between log(brain) and body weight could be improved by a log transformation of body weight.



```{r}
#pairwise comparison of log(brain) and log(body)#
plot(log(animals$Brain)~log(animals$Body))


#Full Model#
brain.lm<-lm(log(Brain)~log(animals$Body)+animals$Gestation+animals$Litter,data=animals)


summary(brain.lm)

#Checking MLR assumptions#

#Resid vs Fitted#
plot(brain.lm,which=c(1))

#Normality of Residuals#
hist(brain.lm$residuals)

#Cooks Distance#
plot(brain.lm,which=c(4))

#leverage#
lev<-hat(model.matrix(brain.lm))

plot(lev)

#Added Variable Plots#
library(car)
avPlots(brain.lm)


#Leaps Function#

#log transforming data so that leaps is working with suitable predictor variables#

animals$log.brain<-log(animals$Brain)
animals$log.body<-log(animals$Body)


library(leaps)

leaps(x=animals[,4:6],y=animals[,7],names=names(animals)[4:6],int=T,method="Cp")

```

The model with the lowest Cp statistic (4.0) is the one that includes all three predictor variables. However, the more parsimonious fourth model also has a low Cp statistic (6.7) which includes only Body and Gestation predictors. The following analysis will examine these two models.

```{r}
#Checking MLR Assumptions of 4th Model#

#Fit model with only Body and Gestation#
brain2.lm<-lm(log(Brain)~log(animals$Body)+animals$Gestation,data=animals)

summary(brain2.lm)

#Resid vs Fitted#
plot(brain2.lm,which=c(1))

#Normality of Residuals#
hist(brain2.lm$residuals)

#Cooks Distance#
plot(brain2.lm,which=c(4))

#leverage#
lev<-hat(model.matrix(brain2.lm))

plot(lev)

#Added Variable Plots#
library(car)
avPlots(brain2.lm)


```

The 4th model appears to satisfy the assumptions of MLR just as well as the full model.  The two predictors in the 4th model: log(body) and Gestation account for 93.8% of the variation in log(brain).  The full model has an R^2 of 94.9 which is not very much additional explanatory power. I would put the 4th model forward as the best model because it satisfies MLR assumptions and requires on less variable while retaining a very high R^2.  




