---
title: 'Lab #13'
author: "Colin Kirkmire"
date: "May 8, 2016"
output: word_document
---

HW[1]: A company specializing in “spam” mail wants to know what attracts customers to open their spam mail. They send 5000 pieces of mail out, with sometimes using the
first NAME on the envelope and sometimes not, as well as sometimes describing an OFFER on the envelope face, and sometimes not. Results are shown in the table below,
along with the spam.csv file which describes the data.

We want to test the logistic model
logit(OPENED)=β0 + β1 FIRST+ β2OFFER + interactions(if any)

Run a study to see if any of these things get customers to open the mail.

```{r}

spam <- read.csv("G:/STAT 458/Lab13/spam.csv")

spam$unopened<-spam$total-spam$opened

spam$offer<-as.factor(spam$offer)
spam$name<-as.factor(spam$name)

#Interaction Plot for Name and Offer#
library(ggplot2)

ggplot(spam, aes(y=opened,x=name, color=factor(offer)))+
  geom_point(size=3)+
  ggtitle("Number Opened vs Name on Envelope")+
  xlab("Name on Envelope")+
  ylab("Number of Customers that Opened")+
  labs(color="Offer")+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14))
```


An additive model appears to be appropriate based on the interaction plot above that reveals no interaction between whether the company includes the name and an offer in the spam mail.


```{r}

spam$unopened<-spam$total-spam$opened

no_int<-glm(formula=cbind(spam$opened,spam$unopened)~ spam$name*spam$offer, family=binomial(logit))

int_model<-glm(formula=cbind(spam$opened,spam$unopened)~ spam$name*spam$offer, family=binomial(logit))

no_int_sum<-summary(no_int)
no_int_sum

summary(int_model)

anova(no_int,int_model,test="Chisq")

```

Since residual deviance does not decrease significantly (p=.39) according to the chi-squared test, it does not appear to be neccesary to include the interaction between the covariables. In addition, the interaction term has a large p-value associated with it in the summary of the interaction model.  This corresponds well to the initial interaction plot.


```{r}

est_coefs<-no_int_sum$coefficients[1:3,1]

tbl<-cbind(confint(no_int,level=.95),est_coefs)


#Odds CI for Name#
exp(tbl[3,1:3])

#Odds CI for Offer#
exp(tbl[3,1:3])

```
The odds of a customer opening the spam mail increases by an estimated 1.4 times (95% CI between .84 and 2.38 when name is included on the envelope and offer is held constant. This indicates some inconclusiveness about the effect of including the name on the envelope beacue the interval conatins values less than 1 and is fairly wide.

In contrast, the odds of a customer opening the spam mail increases by 





```{r}
plot(jitter(Acceptance,factor=.3)~ offer, data=spam, pch=c(1,3)[name],
ylab="Probability of Acceptance", xlab="GPA",
main="Probability of Acceptance vs GPA")
legend("topleft",levels(med_data$Sex),pch=c(1,3),lty=c(1,2))
curve(predict(add_log_model,newdata=data.frame(GPA=x,Sex="F",MCAT=36),type="response"),add=T)
curve(predict(add_log_model,newdata=data.frame(GPA=x,Sex="M",MCAT=36),type="response"),lty=2,add=T)
```
















