---
title: "kirmire_HW4"
author: "Colin Kirkmire"
date: "March 18, 2016"
output: html_document
---


1)
(a) Create a scatterplot of log(Force) vs. log(Height) with Species coded by color (include a
legend). Does the relationship between the variables appear similar for the three species?


```{r}
 Crabs <- read.csv("F:/STAT452/Homework#4/Crabs.csv")
 library(ggplot2)
 
 Crabs$logForce<-log(Crabs$Force)
 Crabs$logHeight<-log(Crabs$Height)
 Crabs$Species<-as.factor(Crabs$Species)
 
ggplot(Crabs, aes(x=Crabs$logForce,y=Crabs$logHeight,color=Crabs$Species))+
  geom_point(size=2)+
  geom_smooth(method="lm",se=FALSE)+
<<<<<<< HEAD
  ggtitle("Log(Height) vs Log (Force) by Species")+
=======
  ggtitle("Log(Height) vs Log (Force) by Factor")+
>>>>>>> 5a516aa4415a2a3738602bd49690c54ef9ddbc65
  xlab("Log(Force)")+
  ylab("Log(Height)")
  
```

(b) Fit a model which assumes a linear relationship between log(Force) and log(Height)
with different intercepts and different slopes for the three species. Assess the model
assumptions.

```{r}
<<<<<<< HEAD
full.lm<-lm(Crabs$logHeight~Crabs$logForce*Crabs$Species)

summary(full.lm)

#Residual Plots#
boxplot(resid(full.lm)~predict(full.lm))
hist(resid(full.lm),col="gray")
qqnorm(resid(full.lm))
=======
model1<-lm(Crabs$logHeight~Crabs$logForce+Crabs$Species+Crabs$logForce:Crabs$Species)
>>>>>>> 5a516aa4415a2a3738602bd49690c54ef9ddbc65

```



(c) Test the hypothesis that all three species have the same slope with an F test.
```{r}
reduced.lm<-lm(Crabs$logHeight~Crabs$logForce)

anova(reduced.lm,full.lm)

```

(d) Test separately (with t-tests) the hypothesis that C. productus and H. nudus have the
same slope. Calculate and interpret a 95% confidence interval for the difference in the
slopes.

```{r}


```

(e) Repeat part (d) for C. productus and L. bellus.

```{r}


```
2. The data file Real estate project.csv contains sales data on 894 houses: price ($), living
area (sq. ft), number of bedrooms, number of bathrooms, year of construction, and garage
spaces.

• Row 185 contains a missing value for bathrooms; just eliminate the row at the start. If
df is the data frame, then do df <- df[-185,].

• Compute Age = 2012-Year (e.g., df$Age <- 2012-df$Year). You can them eliminate
the Year variable with df$Year <- NULL.

• Present a scatterplot matrix (I suggest using ggpairs in library GGally). Do you notice
any strong associations with Price? Are any transformations suggested?

```{r}
real.estate <- read.csv("F:/STAT452/Homework#4/Real_estate_project.csv")

real.estate<- real.estate[-185,]

real.estate$age<-2012-real.estate$Year

real.estate$Year<-NULL

library(GGally)

pa<-ggpairs(real.estate)

```

• Fit a model for predicting Price from all the other variables. Note: a shorthand way to
specify such a model is lm(Price∼.,data=df. The . means “all the other variables in
the data frame”. Assess the model assumptions with residual plots. Do they appear to
be met?

```{r}
full.lm.real.est<-lm(Price~.,data=real.estate)

#Residual Plots#
boxplot(resid(full.lm.real.est)~predict(full.lm.real.est))
hist(resid(full.lm.real.est),col="gray")
qqnorm(resid(full.lm.real.est))
```


1
• Fit the model using log of Price as the response variable. Are the model assumptions
better satisfied?

```{r}
logP.full.lm.real.est<-lm(log(Price)~.,data=real.estate)

summary(logP.full.lm.real.est)
#Residual Plots#
boxplot(resid(logP.full.lm.real.est)~predict(logP.full.lm.real.est))
hist(resid(logP.full.lm.real.est),col="gray")
qqnorm(resid(logP.full.lm.real.est))
```


• Garage and Age both have big P-values. Could we reasonably eliminate both? Test
the hypothesis that both coefficients are 0 with an “Extra Sum of Squares” F test.

```{r}
reducedre.lm<-lm(log(Price)~Living.Area+Bedrooms+Bathrooms,data=real.estate)

anova(reducedre.lm,logP.full.lm.real.est)

```

• Fit the model without Garage and Age. What happens to se? Does this model appear
about as good as the model with all the variables?

```{r}
summary(reducedre.lm)

#Residual Plots#
boxplot(resid(reducedre.lm)~predict(reducedre.lm))
hist(resid(reducedre.lm),col="gray")
qqnorm(resid(reducedre.lm))
```


• Interpret the coefficients of the reduced model. Interpret the coefficient for Living Area
in a meaningful increment. That is, rather than saying what change in Price a 1 square
foot increase in area is associated with, say what a 100 sq. ft. increase in area would do.
Also, remember, it’s a log Price so interpret the coefficient in terms of percent increase
in Price.



• Find confidence intervals for the coefficients and interpret them (you can integrate this
into the previous part).
